---
title:    "CASE STUDY OF PRESSURE MANAGEMENT"
subtitle: "System design in EPANET"
author:   "Dr. Raúl Trujillo Álvarez"
date:     "`r format(Sys.time(), '%d %B %Y')`"

output: 
  html_document:
    theme: spacelab
    toc: true
    number_sections: true

params:
  net_work: "prv_01"

---


```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE)

# Initialize Session
  cat("\014")

# Installs libraries 

  list.of.packages <- c("epanet2toolkit","epanetReader",
                        "data.table","tidyverse",
                        "knitr","kableExtra","captioner") 
  
  new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
  if(length(new.packages)) install.packages(new.packages,dependencies=TRUE)
  lapply(list.of.packages, require, character.only = TRUE)
  rm(list.of.packages,new.packages) # remove 


# initialize files paths and files

  dir_work   <-  gsub("/reports","",getwd())
  dir_report <-  file.path(dir_work,"reports") 
  dir_data   <-  file.path(dir_work,"data")  
  dir_bin    <-  file.path(dir_work,"reports")

# Tables and figures numbering
# R Markdown: How to number and reference tables 
# https://www.r-bloggers.com/r-markdown-how-to-number-and-reference-tables/
  
  nums_tables <- captioner::captioner(prefix = "Tab.")
  nums_fig    <- captioner::captioner(prefix = "Fig.")

  fig.001 <- nums_fig(name = "net_charact",
                      caption = "Scheme of the network fed by deposit")


# The input processor module receives a description of the network 
# being simulated from an external input file (*.INP)


  file_inp     <- file.path(dir_data,   paste(params$net_work,".inp", sep=""))
  file_report  <- file.path(dir_report, paste(params$net_work,".rpt", sep=""))
  file_bin     <- file.path(dir_report, paste(params$net_work,".bin", sep=""))
  
  net_input_01  <- read.inp(file_inp)
  
  entries       <- summary(net_input_01)

```

# DESCRIPTION 

This first analysis is based on a simple water distribution systems, 
which could represent the drinking water supply scheme of a small population,
formed by `r entries$entryCounts[1,1]`  knots and `r entries$entryCounts[3,1]` 
pipes lines, whose characteristics are contained in `r nums_fig("net_charact")`:

```{r network_characteristics, fig.cap = fig.001, echo=FALSE}
  plot(net_input_01)
```

```{r echo=FALSE, results='asis'}
# boot_options <- c("striped", "hover", "condensed", "responsive")
# entries$entryCounts %>%
#  kable("html", caption = "Summary") %>%
#  kable_styling(bootstrap_options = boot_options, 
#                full_width = F, 
#                font_size = 12,
#                position = "left")


```

Two types of demends are considered: consumptions and leakage.


## Time-Varying Demands (consumptions)

All the distribution systems is inherently unsteady due to continuously varying
demands. 

In order for an extended period simulation to accurately reflect the dynamics 
of the real system, these demand fluctuations must be incorporated into the 
model.

The temporal variations of flow (consumption) for water systems follow a 24-hour 
pattern _(diurnal demand pattern)_.

These Flow-Pattern not only are changing on a daily basis, also are different 
in the weekdays (Weekend usage patterns often differ from weekday patterns),
change weekly (vacation, christmas, ramadan etc.) annually (summer and winter) 
and from year to year ( _"El Niño"_ and _"La Niña"_ phenomenon).

Normally, the hydraulic models express demands by using a constant baseline
demand multiplied by a dimensionless demand pattern factor at each time 
increment. 

A demand multiplier is defined as:

For this exercise we are going to limit ourselves to a single daily pattern 
multipliers $(m_{t})$.

$$ m_{t} = \frac{q_{t}}{q_{base}}  $$

where: 

 - $m_t =$ demand multiplier at the $t^{th}$ time step;
 
 - $q_{t}=$ demand in $t^{th}$ time step $(l/s)$  
 
 - $q_{base}=$ base demand $(l/s)$


## Water Demand

selecting $150 lpcd$ (_litres per capita per day water supply_ including 15% for 
unaccounted for flow). 

The design of the hydraulic system is performed for the daily mean
water demand, which total value is $q_m = 39.13 \, (m^3/h) = 10.869 \,(l/s)$ 

## Pressure dependent demands (leaks)

Simulations are run for two types of scenarios: with and without leak. Scenario 
without leak uses the information of consumption and border conditions of a 
standard day (or other time period, if necessary), while scenarios with leak 
incorporates additional node pressure-dependent flow.

Leaks are incorporated in the simulation using nodal emitters whose flow 
value $q_{l_{t}}$ follows the _Equation:_

$q^{leak}_k =(\beta_k*l_k)*p^{\alpha}_k+C_k*p^{\delta}_k$

where: 

 - $q^{leak}_k$ is the leak flow in along pipe $k$ at the $t^{th}$ time step,
 
 - $l_k$ is the length of pipe $k$
 
 - $p_k=$  is the average pressure in pipe $k$ computed as the mean of the pressure values of its end nodes at the  $t^{th}$ time step,
 
 - $\alpha_k=$ is the exponent dependent on pipe and leak typ, in these case we will use $α = 1.18$. The $\alpha_k$ parameter can take values between 0.5 and 2.5, depending on the leak type.
 
 - The $\beta_k$ parameter is related to the pipe material deterioration and its value must be set by calibration (initial value can be set around $0.05$)
   
 - $C_k$ and $\delta_k$ are parameter of the _bursts leakage_ model (classical orifice flow formulas)
 
 For each pipe, the total leakage is assigned to its end nodes, half to each node. So, the nodal leakage flow $q^{leak}_i$ for a node $i$ can be computed as follows:
 
 $q^{leak}_i=\frac{1}{2}*\sum_{k=1}^{n} q^{leak}_k$
 
Where $k$ interates over all $n$ pipes connected to node $i$.



