---
title:    "CASE STUDY OF PRESSURE MANAGEMENT"
subtitle: "System design in EPANET"
author:   "Dr. Raúl Trujillo Álvarez"
date:     "`r format(Sys.time(), '%d %B %Y')`"

output: 
  html_document:
    theme: spacelab
    toc: true
    number_sections: true

---

```{r setup, include=FALSE}

# Initialize Session

  cat("\014")
  rm(list=ls()) 

# Installs libraries 

  list.of.packages <- c("epanet2toolkit","epanetReader",
                        "data.table","tidyverse",
                        "knitr","kableExtra","captioner") 
  
  new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
  if(length(new.packages)) install.packages(new.packages,dependencies=TRUE)
  lapply(list.of.packages, require, character.only = TRUE)
  rm(list.of.packages,new.packages) # remove 


# initialize files paths and files

  knitr::opts_chunk$set(echo = TRUE)

  dir_work   <-  gsub("/reports","",getwd())
  dir_report <-  file.path(dir_work,"reports") 
  dir_data   <-  file.path(dir_work,"data")  
  dir_bin    <-  file.path(dir_work,"reports")

# Tables and figures numbering
# R Markdown: How to number and reference tables 
# https://www.r-bloggers.com/r-markdown-how-to-number-and-reference-tables/
  
  nums_tables <- captioner::captioner(prefix = "Tab.")
  nums_fig    <- captioner::captioner(prefix = "Fig.")

  fig.001 <- nums_fig(name = "net_charact",
                      caption = "Scheme of the network fed by deposit")


# The input processor module receives a description of the network 
# being simulated from an external input file (*.INP)

  net_name     <- "prv_01" #!!!!!!!
  
  file_inp     <- file.path(dir_data,   paste(net_name,".inp", sep=""))
  file_report  <- file.path(dir_report, paste(net_name,".rpt", sep=""))
  file_bin     <- file.path(dir_report, paste(net_name,".bin", sep=""))
  
  net_input_01  <- read.inp(file_inp)
  
  entries       <- summary(net_input_01)

```

# DESCRIPTION

This first analysis is based on a simple water distribution systems, 
which could represent the drinking water supply scheme of a small population,
formed by `r entries$entryCounts[1,1]`  knots and `r entries$entryCounts[3,1]` 
pipes lines, whose characteristics are contained in `r nums_fig("net_charact")`:

```{r network_characteristics, fig.cap = fig.001, echo=FALSE}
  plot(net_input_01)
```

```{r echo=FALSE, results='asis'}
# boot_options <- c("striped", "hover", "condensed", "responsive")
# entries$entryCounts %>%
#  kable("html", caption = "Summary") %>%
#  kable_styling(bootstrap_options = boot_options, 
#                full_width = F, 
#                font_size = 12,
#                position = "left")


```

## Time-Varying Demands

All the distribution systems is inherently unsteady due to continuously varying
demands. 

In order for an extended period simulation to accurately reflect the dynamics 
of the real system, these demand fluctuations must be incorporated into the 
model.

The temporal variations of flow (consumption) for water systems follow a 24-hour 
pattern _(diurnal demand pattern)_.

These Flow-Pattern not only are changing on a daily basis, also are different 
in the weekdays (Weekend usage patterns often differ from weekday patterns),
change weekly (vacation, christmas, ramadan etc.) annually (summer and winter) 
and from year to year ( _"El Niño"_ and _"La Niña"_ phenomenon).

Normally, the hydraulic models express demands by using a constant baseline
demand multiplied by a dimensionless demand pattern factor at each time 
increment. 

A demand multiplier is defined as:

For this exercise we are going to limit ourselves to a single daily pattern 
multipliers.
